### 2017-01-20 intro.tex
Made Christine's revisions

\section{The Cosmological Context of Galaxy Formation}
The modern picture of galaxy formation sits within the context of a $\Lambda$
Cold Dark Matter  ($\Lambda CDM$) cosmology
\citep{Rees1977,White1978,Blumenthal1984,Mo1998}.  This is a universe which has
a no curvature ($\Omega = 1$), and today has the majority of its mass-energy in
a cosmological constant $\Lambda$ ($\Omega_\Lambda \sim 0.7$), and most of its
matter in a collisionless form (dark matter) ($\Omega_m \sim 0.3$)
\citep{Planck2014}.  Roughly $5\%$ of the matter-energy density of the universe
exists as baryons ($\Omega_b \sim 0.05$), which existed as a nearly isothermal
($\delta T/T\sim10^{-5}$) gas of hydrogen and helium at the earliest
observational time.  The composition of the universe, and the precise value of
the mass-energy density for these components sets the lifetime of the universe,
how rapidly it expands (or contracts), and has tremendous impact on the growth
of structure.  This structure is ultimately where galaxies will form. All of
this began $13.8$ billion years ago, with the universe emerging from a state of
high temperature and density, expanding to form what we see today.  The
primordial elemental abundances, with $\sim4/5$ of the universe's baryons in
hydrogen, and the remainder in helium, was established just a few minutes after
the big bang, during a brief period of ``Big Bang Nucleosynthesis''
\citep{Alpher1948}.  The remaining heavier elements have been formed via stellar
nucleosynthesis, through fusion in the cores of stars and in energetic processes
that take place during supernovae (SN) \citep{Wagoner1967}, or through the
capture of high energy cosmic ray protons and $\alpha$ particles by carbon, nitrogen, or
oxygen nucleii \citep{Reeves1970}.  The chemical composition of the universe
establishes the kinds of stars that form within galaxies and the rates at which
gas within the ISM and beyond cools.

Much of what we know about the cosmology of our universe comes from observations
of the first light to escape the universe's surface of last scattering as it
became optically thin: the cosmic microwave background radiation (CMB).  Prior
to a redshift of $z\approx1100$, the universe had a high ionization fraction
such that the mean free path for photons was short, making all of space
effectively opaque.  Once the universe became cool enough that hydrogen became
fully neutral, the optical depth fell low enough that photons could free-stream,
decoupled from baryons.  This ultraviolet light we observe today, redshifted to
microwaves with a temperature of $\sim2.7\;\rm{K}$.  When the power spectrum of
this radiation is measured \citep{Spergel2003}, features in that spectrum can be
used to determine the components of the universe (with some ambiguity that can
be resolved via SN Ia and large scale structure measurements
\citealt{Riess1998,Perlmutter1999,Beutler2011,Blake2011}), as well as the
spectrum of density perturbations that will eventually form galaxies
\citep{Press1974,Peebles1980}.   The CMB is the single most important
observational tool for understanding the cosmology of our universe.

The formation of galaxies within the universe began with the gravitational
collapse of small, linear density perturbations that were likely seeded by
quantum fluctuations amplified by inflation \citep{Guth1981,Linde1982}.  These
perturbations grow through gravitational collapse.  Until $z\sim1100$, only dark
matter perturbations were able to grow.  Prior to this, baryons were coupled to
photons, causing density perturbations to oscillate as stable sound waves.  This
early collapse of matter meant that when decoupling occurred at $z\sim1100$,
baryons were able to collapse into the deeper potential wells grown out of
dark-matter density perturbations.  These overdensities expand slower that the
surrounding medium, as they experience a drag from their own self-gravity.
Eventually, these regions stop expanding altogether and begin collapsing at the
turnaround time $t_{turn}$, and virialize at $2\;t_{turn}$ \citep{Peebles1980}.
The first galaxies likely began to form around $z\sim20-50$.  Without the
earlier growth of dark matter perturbations, structure formation would have
occurred much later, and the number of galaxies and distribution of their masses
would be different than those we see today \citep{Davis1985}.

The perturbation spectrum that forms the structure in which galaxies begins to
form is essentially a Gaussian random field, which \citet{Press1974} were able
to show produces a halo mass spectrum at a given time with a form that matched
the functional form for the observed galaxy mass function
\citet{Schechter1976}:
\begin{equation}
    \frac{dn}{dM}(M,t) =
    \sqrt{\frac{2}{\pi}}\frac{\rho_0\delta_c(t)}{M^2\sigma(M)} 
    \left\lvert\frac{d\ln{\sigma}}{d\ln{M}}\right\rvert
    \exp{\left[-\frac{\delta^2_c(t)}{2\sigma^2(M)}\right]}
\end{equation}
In this equation, $\rho_0$ is the mean density of the universe, and
$\delta_c(t)$ is the critical overdensity for collapse at a given time $t$.
The mass variance spectrum, $\sigma^2(M)$ can be determined observationally
using the CMB power spectrum \citep{Fixsen1996,Spergel2003,Planck2015} and an
assumption of a specific cosmology.  This function gives a distribution of
masses for the dark matter halos that galaxies will begin to form within.  As
these halos collapse, gravitational tugs from their surroundings will impart
torques on them \citep{Barnes1987}, causing halos to have a distribution of
angular momenta.  High resolution N-body simulations have found that dark
matter halos have both a universal density profile
\citep{Navarro1996,Merritt2006} as well as a universal angular momentum profile
\citep{Bullock2001}.  These dark matter halos all serve as the environment in
which the collapse of baryons and formation of stars take place.

The collapse of baryons within halos will depend strongly on whether the
post-virialized gas can cool radiatively \citep{Rees1977}.  As gas accretes onto
a halo with overdensity $\Delta$ it passes through a shock that raises its
temperature to the virial temperature:
\begin{equation}
    T_{virial} = \frac{2}{3}\frac{G\mu m_H}{k_B}\left({\frac{4\pi\rho_0\Delta
    M^2}{3}}\right)^{1/3}
\end{equation}
This means that more massive halos have larger virial temperatures, and,
depending on the mode of radiative cooling (atomic cooling from primordial
hydrogen, molecular hydrogen cooling, or cooling through metal lines), this may
give cooling times shorter than the current age of the universe.  These halos
are the ones in which galaxy formation may begin, as their hot gaseous halos can
collapse to form a central concentration where star formation can occur.  In
some halos, these virial shocks may be unstable to cooling, allowing the direct
collapse of cold gas onto the disc \citep{Birnboim2003}.

As gas does begin to cool within these halos, it will lose pressure support (by
radiating away thermal energy), but {\it not} angular momentum.  For large
spiral galaxies, this means that the baryons, which typically carry the same
specific angular momentum as the halo itself, will flatten to form an
exponential disc as they cool \citep{Fall1980,Mo1998}.  Streams of gas
accreting along the cosmic web may also provide angular momentum to the forming
disc \citep{Dekel2006}.  Within this disc, an ISM will form.  This ISM is
multiphase, containing both hot ($T\sim10^4\;\rm{K}$) and cold
($T\sim10-100\;\rm{K}$) gas in quasistatic equilibrium \citep{McKee1977}.  This
multiphase state is highly dependent on both the ISM metallicity and the local
heating rates (which themselves can vary with redshift)
\citep{Field1969,Wolfire1995,Norman1997}.
\begin{figure}
    \includegraphics[width=0.7\textwidth]{mass_function.eps}
    \caption[Galaxy mass function]{The observed mass function for galaxies in
    the nearby universe (solid line) does not match the mass function for dark
    matter halos given in equation 1, scaled by the baryon fraction (dashed
    line).  This tells us that there are additional physical processes involved
    in the formation of galaxies within their dark matter halos.  The most
    likely candidate for this is feedback from stars and black holes.  Of
    particular interest is the peak in star formation efficiency near
    $M_{200}=10^{12}M_\odot$.  \textit{Image credit: Adapted from figure 1 in
    \citet{Ferrero2012}}}.
\end{figure}

These are of course all highly idealized, simple models for the broadest strokes
of galaxy formation.  Each of these halos will begin to undergo mergers and
interactions \citet{Navarro1993,Kauffmann1999,Cole2000} with each other that can
strip stars and gas through tidal and ram pressure forces
\citep{Gunn1972,Dressler1980}.  Inside each of these halos, the process of disc
collapse, the formation of a multiphase ISM, star formation, and feedback from
those stars involves the complex, nonlinear interplay of radiation,
hydrodynamics, magnetic fields, and gravity on length scales as much as
$10^{11}$ times smaller than the virial radius.  None of this physics is simple
enough to be fully captured by analytic models. As figure 1.1 shows, the
observed mass function for galaxies does not match what would be expected from
the total baryon budget for each halo, implying that at least some of these
processes must change the galactic star formation efficiency as a function of
halo mass, establishing the population of galaxies we see today.  Numerical
simulations have therefore become a keystone for galaxy formation research.

\section{Numerical Simulations of Galaxies}
N-body Simulations of galaxies actually predate {\it digital} simulations of
galaxies.  \citet{Holmberg1941} presented, as the author termed it, a ``new
integration procedure'' that was the first attempt to simulate the
evolution of a galaxy.  By relying on the fact that light flux, just as gravity,
scales as $r^{-2}$, Holmberg was able to construct a simulated galaxy disc out
of light bulbs, and then using photocells to measure the flux at each bulb's
position, repositioned the bulbs using the forces ``simulated'' by the
light-measuring process.  This allowed him to painstakingly trace the tidal
interactions of passing galaxies as they merged with each other.

In the intervening 75 years, we have migrated to much less labor-intensive
simulations, relying on the constantly growing power of digital computers.  As
Gordon Moore first observed in 1965, the processing power of computers has been
continuously growing at an exponential rate for more than half a century.  This
has meant that, even without advances in numerical methods and algorithm design,
the ability of simulators today to model a system in high resolution is
monumentally greater than just 10 years ago.

\begin{figure}
    \includegraphics[width=0.8\textwidth]{Toomre.eps}
    \includegraphics[width=0.8\textwidth]{Antennae.eps}
    \caption[Early simulation of Antennae Galaxies]{The simulations of
    \citet{Toomre1972} showed how the kinematics of galaxy mergers can produce
    the variety of tidal structures seen in objects like the Antennae galaxies.
    The top image shows the result of one of the \citet{Toomre1972} simulations.
    \textit{Image credit: Adapted from figure 23 of \citet{Toomre1972}}
    The bottom image shows an observation of the actual object itself, the
    Antennae galaxies NGC 4038/4039. \textit{Image credit: NOAO/AURA/NSF, B.
    Twardy, B. Twardy, and A. Block (NOAO)}}
\end{figure}

The earliest numerical simulations of galaxy formation focused on the effects of
gravity alone.  A classic example of this are the simulations presented in
\citet{Toomre1972}, which showed that the irregular structures seen in galaxies
like NGC 4676 (``The Mice'') or NGC4038/4039 (``The Antennae Galaxies'') are the
results of major mergers between two disc galaxies, as is clear from figure 1.2.
These simulations used discs composed of a mere 120 test particles, and ignored
the effects of self-gravity between these particles.  Despite their crudeness,
these simulations definitively showed that the morphologies of galactic bridges and tails could be
formed by the tidal interactions of two passing/merging galaxies.  The IBM
360/95 that \citet{Toomre1972} (with an inflation-adjusted purchase price in the
millions of dollars) used for their simulations was capable of 5.5 million
instructions per second, roughly $0.2\%$ the speed of a 35\$ Raspberry Pi 3
available today.

Modern galaxy simulations do far more than merely integrate the equations of
motion for a few hundred test particles in a gravitational field.  Today, state
of the art simulations will track gravity (including self-gravity) for stars,
gas, and dark matter \citep{Aarseth1980,Stadel2001,Dehnen2002}; evolve Euler's
equations for the hydrodynamics of gaseous baryons
\citep{Wadsley2004,Teyssier2002,Bryan2014}; include models for radiative
cooling of gas that allow for cooling from both hydrogen and metal lines
\citep{Marri2003,Shen2010}; form stars from dense gas
\citep{Katz1992,Agertz2015}; and include models for feedback from massive stars
and black holes \citep{Katz1992,Springel2003,DiMatteo2005}.  Often, simulations
will also include treatments for radiative transfer
\citep{Gendelev2012,Krumholz2013} and magneto-hydrodynamics
\citep{Girichidis2015} as well.

Numerical modeling of galaxies is a difficult problem, as it involves processes
that take place over scales (both spatial and temporal) that span many orders of
magnitude.  This means that simulations will always be unable to resolve at
least some fraction of the physics taking place within the galaxy.  For this,
carefully designed ``sub-grid'' models are required to include the macroscopic
effects that unresolvable microscopic processes yield.

Gravity is typically solved using algorithms designed to minimize the workload
that would be needed for a direct summation of each particle or grid cell
($\mathcal{O}(N^2)$ for N grid cells/particles).  This becomes intractable for
larger numbers of resolution elements, as the workload will scale as the sixth
power of the linear resolution (an order of magnitude better linear resolution
will be $10^6$ times more expensive!).  In order to reduce these costs,
algorithms have been developed that allow for more efficient gravitational
force calculations.  Tree-based methods \citep{Barnes1986} reduce the
computational cost of long-ranged (and thus less significant) force
calculations by using space-partitioning trees.  Short-range force calculations
are done particle-by-particle, while long-range forces are calculated against
larger tree cells, thus reducing the cost of calculating gravity to
$\mathcal{O}(N\log N)$.  Particle-Particle/Particle-Mesh ($P^3M$) methods
\citep{Couchman1991} produce similar scalings, but rather than using a
spatial tree rely on a spectral Fourier-based approach to calculate long-ranged
interactions.  Recently, the Fast Multipole Method \citep{Greengard1987} has
been applied to the calculation of gravitational forces
\citep{Dehnen2002,Hahn2013} using a multipole expansion of Green's functions.
This allows for extremely efficient gravity calculation, scaling linearly with
particle number ($\mathcal{O}(N)$).

Hydrodynamical solvers usually come in one of two classes: Eulerian or
Lagrangian.  Eulerian methods discretize space into fixed volume elements with
static boundaries.  By breaking a simulation volume into a grid of cells that
each contains scalar and vector quantities those quantities can be advected
between cells according to the solver scheme.  Eulerian schemes typically allow
for higher-order solvers, lower intrinsic noise, and better resolved
discontinuities \citep{Teyssier2002,Stone2008,Bryan2014}.  They suffer from
intrinsic mixing and diffusion that scales with the flow velocity relative to
the grid, poor integration of circular orbits, and costly timestep limits
\citep{Agertz2007,Tasker2008}.  Lagrangian methods instead discretize mass,
breaking the gaseous volume into discrete points, each of which moves at the
fluid velocity to advect the quantities it contains, with smoothing functions
used to interpolate values in between particles; hence the name for the most
common Lagrangian method, Smoothed Particle Hydrodynamics (SPH).  Lagrangian
methods couple easily to tree codes for gravity calculation, are Galilean
invariant, and intrinsically conserve mass, energy, and momentum (linear and
angular) \citep{Katz1996,Wadsley2004,Springel2005}.  SPH requires the use of
artificial viscosity to handle shocks, and as such have difficulty in capturing
subsonic turbulence \citep{Bauer2012}. Recently, new methods have begin to
hybridize features of each of these classes as well
\citep{Springel2010,Hopkins2015}.

Modern galaxy simulations also now use a variety of strategies for judiciously
applying computational power where it is most needed.  Eulerian methods will
commonly use \textit{Adaptive Mesh Refinement} to increase resolution in regions
of interest (typically where gas is densest).  Lagrangian methods automatically
do this, as a denser region definitionally must contain more fluid elements.

Beyond the choice of algorithms, the choice of what type of simulation to set up
in the initial conditions is also important.  To study large populations of
galaxies, along with large-scale structure, large-volume cosmological boxes with
volumes $\gtrsim 10^6\;rm{Mpc^3}$, but with spatial resolutions $\sim \rm{kpc}$
are used.  Studying how individual galaxies form is typically done with
cosmological zoom simulations, where a comparatively cheap dark-matter only
simulation is first run to select a region of interest, which is then
resimulated with higher resolution and baryonic physics included
\citep{Navarro1993}. Zoom-in simulations can typically achieve an order of
magnitude or better resolution than full cosmological boxes.  Isolated galaxy
disc simulations cannot show how galaxies form in a cosmological environment,
but can be run with $\sim \rm{pc}$ resolution to study closely how internal
processes within a galaxy take place \citep{Hopkins2011,Benincasa2016}.  These
details of these internal processes are discussed in section 1.3.

The earliest attempt to track the formation of a galaxy in a cosmological
context with the effects of star formation and feedback included was by
\citet{Katz1992}.  This simulation used the Lagrangian TREESPH method
\citep{Hernquist1989} to evolve a parcel of gas and dark matter from an initial
linear perturbation, including small-scale power derived from CDM predictions
\citep{Zeldovich1970,Peebles1982}.  Gas was allowed to cool radiatively using
rates calculated for primordial gas.  This study introduced a model for star
formation that is still the standard for most simulations of galaxy formation.
Observations \citep{Kennicutt1998} have suggested that the star formation rate
is a power law of the gas density: $\dot\Sigma_*\propto\Sigma_{gas}^\alpha$, with
$\alpha=1-2$.  If it is assumed that gas collapses to form stars in some
multiple $c_*$ of its freefall time \citep{Schmidt1959}, $t_{ff}$, a simple
3-dimensional star formation model can be built with the form:
\begin{equation}
    \dot\rho_* = \frac{c_*\rho_{gas}}{t_{ff}} = c_*\sqrt{\frac{\rho^3_{gas}}{4\pi G}}
\end{equation}
\citet{Katz1992} used a $c_*$ parameter of $0.1$, which is within the range of
what is typically used in simulations today ($1-100\%$).  On top of the model
for star formation, feedback was included without the use of any significant
subgrid model:  $10^{51}\;\rm{erg/SN}$ worth of energy was injected back into
the ISM by smoothing it over a number of resolution elements.  One of the key
findings of \citet{Katz1992} was that the vast majority of SN energy was
radiated away, making them very ineffective at regulating star formation within
the galaxy.  In the end, \citet{Katz1992} found that the peak star formation in
the simulated galaxy was nearly 10 times greater than in observed galaxies.
\citet{Thacker2000} later showed that the smoothing of energy used by
\citet{Katz1992} resulted in a large overestimation of the cooling rates in
feedback-heated gas. Since then, much work has gone into developing models for
SN feedback that address the issue of overcooling that was present in these
earlier simulations.

\section{The Importance of Feedback}
\begin{figure}
    \includegraphics[width=\textwidth]{M82.eps}
	\caption[Massive outflows in M82]{The evidence for feedback is no clearer
	than here in the Cigar Galaxy, M82.  The Hubble Space Telescope reveals massive
	outflows of hot ionized gas through the red $H\alpha$ emission as gas is
	blasted out of the galaxy in a bipolar flow. \textit{Image credit: NASA, ESA
	and the Hubble Heritage Team STScI/AURA}}
\end{figure}
We now have strong evidence that galaxy and star formation is not a one way process of
gravitational collapse.  A number of internal processes (``feedback'') within
galaxies can liberate energy greater than the gravitational binding energy of
molecular clouds (and indeed even comparable to the binding energy of the entire
halo).  Analytic calculations that omit feedback processes found that gas should
rapidly collapse into the smallest halos at high redshift, overproducing
stars in small galaxies at high redshift (the ``cooling catastrophe'')
\citep{Cole2001,Benson2003}.  Simulations of galaxies like our own develop unrealistically
compact morphologies without feedback \citep{Stinson2006}.  Without feedback, simulations of
individual molecular clouds have star formation rates an order of magnitude
too high \citep{Agertz2013}.  Aside from regulating star formation and the
growth of galaxies, there are also observations of the large-scale impact of
feedback. Lyman $\alpha$ forest observations probing the intergalactic medium
(IGM) have found that the IGM is polluted with metal ions at great distances
from any galaxies \citep{Sargent1988,Songaila1996,Dave1998}.  These metals must have
been ejected by galactic outflows from the ISM of galaxies where they formed.
As figure 1.3 shows, fast, massive outflows have been observed leaving the ISM of
starburst galaxies and quasar host galaxies \citep{Veilleux2005,Werk2014}.

The two primary sources of feedback energy are massive stars and supermassive
black holes (SMBHs).  Massive stars ($M>5M_\odot$) experience short, violent
lives.  During their lives, they will drive fast $v>1000\;\rm{km/s}$ stellar
winds \citep{Weaver1977}, and heat their surrounding gas with UV radiation
\citep{Krumholz2009}.  At the end of their lives, $4-30\;\rm{Myr}$ later, these
stars die in energetic core-collapse SN, liberating $\sim10^{51}\;\rm{erg}$
\citep{Wilson1985,Bethe1990}.  Figure 1.4 shows the total budget of available
energy liberated by a population of stars.  The energy released by a population
of stars with solar metallicity is $\sim 2\times10^{51}\;\rm{erg/M_\odot}$.
Most of this energy is released as photons, over the lifetime of each star.
These will couple poorly to the surrounding ISM, and so have little impact on
disrupting all but the densest gas.  While stellar winds typically have
comparable luminosity to supernovae, they occur for an order of magnitude less
time ($\sim4$ Myr vs. $\sim40$ Myr).  This means that supernovae, despite only
contributing $\sim1\%$ of the total energy released by a stellar population, may
have the greatest impact on galactic scales.  This energy imparts heat and
momentum to the surrounding ISM, and can accelerate cosmic rays in the shocks
that result \citep{Bell1978}.  In galaxies with actively growing SMBHs, the
accretion disc that forms as gas funnels into the black hole can reach
temperatures exceeding $10^9\;\rm{K}$ through viscous heating
\citep{Antonucci1993}.  This makes these accretion discs some of the most
luminous objects in the universe, and can heat the gas that surrounds the galaxy
to choke off the accretion of gas onto the galaxy, shutting off star formation.

\begin{figure}
    \includegraphics[width=0.8\textwidth]{FB_budget.ps}
    \caption[Stellar feedback energy budget]{Most of the energy released by a
    stellar population comes from the luminosity of the stars within that
    population.  Despite this, because supernovae couple more easily to the ISM,
    it is believed they have the greatest effect on galactic scales.  The data
    shown here was produced with {\sc starburst99} \citep{Leitherer1999}}
\end{figure}

The coupling of feedback energy to the ISM is what determines the effectiveness
of that feedback.  In the lowest mass galaxies, UV photoheating may be the
primary regulator of star formation \citep{Efstathiou1992}.    However, UV
radiation has characteristic temperatures too low to drive galactic winds.
Large {\sc Hii} regions, such as 30 Doradus, have observed expansion rates of
only $25\;\rm{km/s}$ \citep{Chu1994}, far below the velocities needed to drive
galaxy-scale outflows.  Radiation pressure on dust grains has been suggested as
a driving mechanism for outflows \citep{Murray2011}.  Simulations of
radiation-driven galactic winds have thus far relied on major approximations
\citep{Roskar2014,Agertz2015} and questionably large infrared optical depths
\citep{Hopkins2014}.  The porosity of the ISM likely means that the vast
majority of UV photons are able to escape without imparting significant energy
or momentum \citep{Dale2005,Walch2012,Krumholz2013}.

On top of releasing just $10\%$ of the energy released by supernovae, stellar
winds are active during the early stages of a massive star's life.  This means
that these winds are released into the dense natal gas that massive star
clusters form in.  Stellar wind shocks within this dense molecular gas cool
efficiently, leading to much of the stellar wind energy being lost to radiation
that escapes the ISM completely.  Simulations of molecular clouds with feedback
from star clusters \citep{Gendelev2012,Rogers2013} have found that most stellar
wind energy is expended disrupting the molecular clouds the clusters live in.
While these winds do not significantly escape the cluster, the channels they
open within the cloud allow $\sim99\%$ of SN energy to escape the dense
molecular environment.

On the galactic scale, feedback regulates star formation in two primary ways.
Energy and momentum deposited into the ISM can heat and disrupt the cool clouds
of gas which form stars \citep{Rogers2013}, and increase the ISM scale height to
limit the formation of cool clouds out of the warm ISM
\citep{Ostriker2010,Benincasa2016}.  If that feedback is vigorous enough, this
gas can actually be ejected from the disc into the circum-galactic medium (CGM)
\citep{Larson1974,Heckman1987,Hopkins2012b}.  This limits star formation by
simply removing the fuel for the process.  These outflows are needed to explain
observations of low star formation efficiencies in massive galaxies
\citep{Behroozi2013,Moster2013}, observations of outflowing gas
\citep{Lynds1963,Heckman1987}, a metal-polluted intergalactic medium
\citep{Shen2010}, and the low bulge fraction in star forming galaxies
\citep{Brook2012}.  In order to see how the local deposition of feedback energy
drives these winds, simulations with realistic models for hot stellar feedback
are required.

\subsection{Numerical Challenges for Hot Feedback}
The greatest difficulty with including the effects of feedback in simulations of
galaxy evolution is the combination of extreme energy gradients in a relatively
small mass/volume.  Without overwhelmingly high resolution, hot feedback gas
will be numerically mixed with cold ISM, often resulting in gas with
temperatures of $\sim10^5\;\rm{K}$, at the peak of the ISM cooling curve where
cooling times are extremely short.  This results in feedback energy being
completely lost due to radiative cooling, as was first seen by \citet{Katz1992}.
Indeed, as \citet{Thacker2000} showed, how feedback energy is deposited into the
ISM has a huge impact on whether that feedback has any effect.  This problem has
been tackled using a few different strategies.

Since overcooling is the problem, many models have attacked this directly by
either temporarily disabling cooling \citep{Thacker2000,Stinson2006} or
depositing feedback energy into a second, nonthermal energy component
\citep{Agertz2013}.  While this does limit radiative losses, it produces gas
that exists in an unphysical region of the $\rho-T$ phase diagram, which may not
have the correct entropy to escape buoyantly.  It also has the potential to
introduce an \textit{undercooling} problem as resolution becomes higher and
feedback-heated gas becomes resolvable.  Alternatively, energy may be injected
not as heat, but as momentum, which naturally does not cool
\citep{Scannapieco2006,DallaVecchia2008,Dubois2008}.  Unfortunately, this adds
the additional complexity of momentum cancellation (how to treat neighbouring
stars/clusters, where two stars are depositing momentum with opposing
directions), and if this momentum results in strong shocks, these shocks will
convert the momentum right back into thermal energy that can radiate away the
feedback energy the model is designed to preserve \citep{Durier2012}.  This
often means kinetic/momentum feedback models must be combined with hydrodynamic
decoupling that renders feedback-heated gas into a form that does not interact
with the surrounding ISM \citep{Springel2003,Vogelsberger2013}.  Naturally, this
makes outflow properties strongly influenced by numerical choices
\citep{DallaVecchia2008}.  A third choice is to stochastically deposit feedback
energy only when enough energy is available to produce gas temperatures high
enough for cooling times to be short \citep{DallaVecchia2012,Crain2015}.  In
these models, SN events are effectively grouped together, somewhat decoupling
feedback spatially from star formation.  The choice of temperature to heat gas
to is a purely numerical parameter, and can dramatically change the
effectiveness of these models.  Finally, some \citep{Springel2003,Murante2015}
attempt to directly address the unphysical numerical mixing of hot and cold gas
by separating the two phases and tracking each component separately.  This
completely solves the overcooling problem, and allows the two components to cool
radiatively.   Unfortunately, it also means that cold star-forming gas is
permanently coupled to hot feedback-heated gas: preventing the simulation of  a
resolved multiphase ISM, and effectively making the ISM an anchor holding down
potential hot outflows.  This means these multiphase models have to be coupled
with parameterized models to capture outflows \citep{Springel2003}.  Often,
these simple models simply choose a fraction of feedback energy to be deposited
as kinetic ``kicks'' to a some gas particles.  Like other kinetic feedback
models, this is frequently combined with hydrodynamical decouplings
\citep{Vogelsberger2013}.

All of these models attempt to solve the problem of overcooling with unphysical
approximations: disabled cooling, hydrodynamic decoupling, stochastic deposition
of feedback energy, and imposed multiphase ISM models.  These choices make it
difficult to understand what details of simulated galaxy evolution are a result
of these numerical free parameters, as opposed to the actual physics of
self-regulated star formation and SN-driven outflows.  What is needed is a model that
captures the effects of feedback while matching as closely as possible the
actual physical processes involved below the resolution scale.

\subsection{Superbubbles: The Natural Scale for Feedback}
The formation of massive stars takes place mostly in clusters of $10^4$ or more
stars.  This means that the SN of these stars can be treated not as
individual dumps of $10^{51}\;\rm{erg/SN}$, but instead as a luminosity of
$10^{34}\;\rm{\frac{erg}{s \cdot M_\odot}}$.  This allows us to use stellar wind
models like \citet{Weaver1977} to study the evolution of superbubbles driven by
clustered SN.  Each individual SN's ejecta thermalizes in a small
region within the center of the bubble, and the resulting hot gas drives a shock
that sweeps up the surrounding ISM.  In a short time, the swept-up ISM can
radiatively cool efficiently , resulting in a thin, cold shell surrounding a
hot, diffuse bubble.

The structure and evolution of superbubbles have been the subject of a great
deal of analytic study.  \citet{Castor1975} found that the time evolution of
superbubbles has a similarity solution, similar to the \citet{Sedov1959}
adiabatic blast wave solution: $R(t)=27n_0^{-1/5}L_{36}^{1/5}t_6^{3/5}\;\rm{pc}$
and $V(t)=16n_0^{-1/5}L_{36}^{1/5}t_6^{-2/5}\;\rm{km/s}$ , where $n_0$ is the
ambient density in $\rm{cm^{-3}}$, $L_{36}$ is the luminosity of the star
cluster in $10^{36}\;\rm{erg/s}$, and $t_6$ is the time in $10^6\;\rm{yr}$.  The
evolution of hot mass in the bubble, and the effects of cooling in the swept-up
shell, were examined in \citet{Weaver1977}.  A major result of this study was
the realization that thermal conduction between the hot bubble and cold shell
will result in an evaporative mass flux, pulling material into the hot bubble.
This evaporated mass will quickly dominate over the small amount of ejecta from
stellar winds.  The \citet{Weaver1977} results were extended to clustered SN by
\citet{MacLow1988}.  They calculated that the typical cooling time for the hot
bubble was $\sim15\;\rm{Myr}$, with a weak dependence on the total luminosity
and ambient density.  Before significant cooling can occur in the hot interior,
$\sim35\%$ of the total feedback luminosity is lost to cooling of the swept-up
shell, which cools in $\sim10^4\;\rm{yr}$.  \citet{MacLow1988} also demonstrated
that, as superbubbles do not have surface tension, they are not depressurized by
cold clouds, and instead simply deform around them while continuing to expand.
\citet{Silich1996} studied the effects of an ISM permeated with cold clumps in
greater detail, finding that the general behaviour discussed in
\citet{MacLow1988} held even in the case of an inhomogeneous ISM.  Even with
$50\%$ of the ISM mass trapped in cold clouds (which penetrate the cold shell of
the superbubble), the total hot mass within the superbubble will only vary by $\sim20\%$.

A key insight in these studies was that hot bubbles would evaporate their cold
shell through thermal conduction.  In a hot ionized plasma, electrons are able
to penetrate temperature gradients and deposit energy within cooler material.
This produces a pressure gradient that establishes a mass flux against the
temperature gradient, evaporating the cold shell into the hot bubble.  This
evaporation means that a hot bubble with temperature $T$ and radius $R$ gains
mass at a rate of:
\begin{equation}
    \frac{{\rm d }M_b}{{\rm d}t} = \frac{16\pi\mu}{25k_B} CT^{5/2}R
\end{equation}
Where $C=6\times10^{-7}\;\rm{erg\;s^{-1}\;cm^{-1}\;K^{7/2}}$, and $\mu$ is the
mean molecular weight of the gas in the bubble.  As the cold, evaporated
material acts to cool the hot bubble, this makes the bubble's interior
temperature self-regulating: higher temperatures give higher mass flux, cooling
the bubble.

\citet{DallaVecchia2012} showed that the temperature of feedback-heated gas can
ultimately determine its effectiveness in driving outflows and regulating star
formation, even with constant feedback energy.  The self-regulating temperature
of superbubbles provides the natural mechanism for setting this temperature:
thermal evaporation.

\section{This Work and Organization of Chapters}
In this thesis, I present a new sub-grid model for the feedback from massive
stars.  I use that model to investigate how this feedback can regulate the
formation of stars within galaxies, and finally show how this same feedback is
unnecessary to produce galaxies that match the observations of
\citet{McGaugh2016}.  I conclude with a summary of these results, their
implications, and some potential future directions for study.

Chapter 2 presents the new superbubble feedback model, and describes the
implementation of this model in the simulation code {\sc GASOLINE2} (Wadsley et
al. 2016, submitted).  I show a series of tests that demonstrate the model is
robust to resolution changes, insensitive to magnetic field strength, and able
to handle cases involving unresolved ISM structure.  Finally, I show that
simulations of isolated dwarf and Milky Way like disc galaxies regulate their
star formation and drive efficient galactic outflows with this new feedback
model.

Chapter 3 applies the feedback model described in chapter 2 to a cosmological
simulation of a Milky Way like galaxy.  This allows us to see how galaxies like
the Milky Way assemble themselves by regulating the flow of gas into the disc,
ejecting material to prevent the formation of a central stellar bulge and slow
the formation of stars.  We compare this to the same simulation run with the
simpler \citet{Stinson2006} feedback model to reproduce the results of
\citet{Stinson2010}, which used the same initial conditions as this study.  We
show that the overproduction of stars seen by \citet{Stinson2010} was not a
failure of SN feedback in general, but that SN feedback simulated without 
thermal evaporation fail to drive mass-loaded winds at high redshift.  These
winds, we demonstrate, remove the fuel for star formation and bulge growth to
produce a realistic, bulgeless disc galaxy.

Chapter 4 extends the results presented in chapter 3 by simulating an additional
17 cosmological galaxies, reproducing the full sample presented in
\citet{Stinson2010} with updated hydrodynamics and the new feedback model
described in chapter 2.  This sample spans the peak of the star formation
efficiency found by abundance matching studies like \citet{Moster2013}, and
using it we are able to show that SN feedback is unable to regulate
galaxies more massive than the peak halo mass of $10^{12}\;M_\odot$.  By
analyzing the outflow properties of these galaxies, we are able to determine a
global relation between the halo mass (or disc mass) and the outflow mass
loadings.  We show that the decreasing efficiency of SN-driven winds is an
inevitable consequence of the energy budget available from SN alone.

Chapter 5 uses the sample of galaxies developed in the previous chapter to
test whether recent observational results presented in \citet{McGaugh2016}
are consistent with $\Lambda CDM$.  The radial acceleration relation (RAR) seen in
the SPARC sample \citep{Lelli2016} of observed rotationally supported galaxies
shows that the accelerations within these galaxies are tightly correlated 
with the accelerations due to baryons alone.  We show, using the sample of
galaxies generated in chapter 4, that this relation simply falls out of the
dissipative collapse of baryons in a collisionless dark matter halo.  We also
show that this result is independent of stellar feedback.  Despite the insensitivity
to feedback processes at $z=0$, we show that feedback does result in evolution
of the RAR with redshift, and predict that a different RAR would be observed if
a similar study to \citet{Lelli2016} were done for high-redshift galaxies.

Finally, in chapter 6 we conclude with a discussion of how each of these results
together improve our understanding of SN-regulated star formation in the
formation of disc galaxies.  We discuss how the results presented here improve
the understanding developed in past research and its impact on the direction of
research in galaxy evolution theory.  We conclude with a discussion of the
future directions that this research points, and a number of open questions
raised by this research.

\bibliographystyle{mnras}
\bibliography{library}