### 2017-04-03 conclusion.tex
Finished personal revisions for conclusion

In this thesis we have explored the effects of SN feedback in the formation of
galaxies like our own.  Supernovae produce extremely hot ejecta
($T>10^8\;\mathrm{K}$) that rapidly shocks and expands against the significantly
under-pressured surroundings.  As massive stars form in clusters, these
individual SN remnants merge together, forming a superbubble.  These
superbubbles are much more effective at pushing and heating the ISM than an
equivalent amount of energy distributed as individual SN.  The primary mechanism
that sets the temperature of these superbubbles is the evaporation of cold gas,
driven by thermal conduction.  Thermal conduction becomes effective above
$10^5\;\mathrm{K}$, and strongly self-limits the temperature of superbubbles to
$\sim2-5\times10^6\;\mathrm{K}$.  In this thesis, we have developed a new model
for SN feedback that takes into account these previously ignored superbubble
processes, and used that model to study the growth and evolution of star-forming
galaxies.  This model allowed us to see how SN can (and cannot!) regulate the
flow of gas into and out of galaxies, where it may forms stars.  Previous work
\citep{Stinson2013,Hopkins2014} has suggested that star-forming galaxies could
only regulate their growth if SN feedback was augmented with early feedback from
radiation pressure, stellar winds, and UV heating.  With careful, physically
motivated feedback models, we have now shown that this is not the case:  the
additional energy \& free parameters provided by early feedback are not
necessary.  The primary insight this thesis provides is that the processes
which mix feedback-heated gas with the ISM (i.e. thermal evaporation) are
just as important as the amount of energy provided by feedback.  Neither can be
omitted if one wishes to construct an accurate simulation of galaxy formation.

In Chapter 2, we presented a detailed description of the new superbubble model,
implemented in the SPH code {\sc Gasoline2}.  We rely on the strongly
self-limiting process of thermal evaporation \citep{Cowie1977}, where cold mass
is evaporated into a hot bubble at a rate strongly dependent on the temperature
of the hot bubble. This drives the interior of the bubble to a temperature of a
few $10^6$ K, evaporating material more or less rapidly as the temperature dips
above or below this value.  This is the characteristic temperature for SN-heated
gas.  We use this physics to develop a new model for SN feedback, which gives
realistic estimates for cooling in unresolved hot bubbles, and accurately
calculates the mass loading of SN energy.  We demonstrated that this model is
insensitive to resolution, unresolved ISM structure, and uncertain (turbulent)
magnetic fields.  The strong temperature dependence of thermal evaporation means
that significant changes to the conduction rate coefficient (such as the suppression by
magnetic fields) result in only slight changes to the equilibrium temperature of
feedback-heated gas.  Finally, after investigating the behaviour of individual
bubbles, we showed that superbubble feedback is more effective at both
regulating star formation and at driving galactic winds in isolated disc and
dwarf galaxies.

Isolated disc simulations, like that shown in chapter 2, lack a full
cosmological history.  To see how SN feedback can impact the overall growth and
history of a galaxy like the Milky Way, we needed to simulate the galaxy from
high redshift to the present.  In chapter 3, we use zoom-in cosmological
simulations \citep{Navarro1993} of a Milky Way-like galaxy to compare the
effects of different feedback models.  The galaxy, when simulated using our new
superbubble model, was able to produce a realistic stellar mass fraction for the
galaxy's entire history.  Not only did the galaxy simulated with superbubble
feedback have the correct stellar mass, we showed that it also was effectively
bulgeless, with a unpeaked, flat rotation curve.  These two properties arose
via a single process.  At high redshift, superbubble feedback efficiently drives
heavily mass-loaded winds, expelling gas from the ISM of the galaxy.  The gas
that forms a stellar bulge is accreted at high redshift, meaning that these
winds preferentially remove low angular momentum material, while preserving the
gas with high angular momentum.  Thus, superbubble feedback is able to naturally
prevent both runaway star formation and the formation of a bulge through these
winds.  Higher-resolution simulations of wind venting will help to determine if
these mass loadings are realistic, and to remove additional uncertainty that
arises from a marginally-resolved ISM.

While a single galaxy can be a useful case study, multiple objects must be
studied in order to look at population-scale effects.  Chapter 4 extends the
results of chapter 3 by simulating an additional 17 objects, and uses these
simulations to examine if SN regulation can operate effectively in galaxies both
heavier and lighter than the Milky Way.  With this new sample (the McMaster
Unbiased Galaxy Simulations 2, or MUGS2), we were able to demonstrate a critical
mass at which SN regulation of gas accretion and star formation breaks down.
Simulated galaxies that exceed this mass $(\sim 10^{12}\Msun)$ overproduce
stars,  most of which form in a dense stellar bulge.  All of this arises due to
a fundamental relation between the mass loading of superbubble-driven winds and
the mass of the galaxy.  When the escape velocity of the disc exceeds the sound
speed of superbubble-heated gas (which has a characteristic temperature of
$\sim3\times10^6$ K), only outflows which have entrained less material are able
to escape the galaxy. This leads to less gas ejection, and a shutdown of
regulation.  This shutdown occurs exactly at the galaxy mass where powerful AGN
are observed, and it results in a vigorous transport of gas to the nucleus of
the galaxy.  This makes these results strong evidence that this shutdown is a key
component in the ``hand-off'' between SN and AGN feedback.

The MUGS2 simulations produced in chapter 4 contain a wealth of data useful for
future studies. Chapter 5 applied the MUGS2 galaxies produced in
\citet{Keller2016a} to examine the radial acceleration relation (RAR) derived by
\citet{McGaugh2016} from observations in the SPARC \citep{Lelli2016b} catalog.
We found that the tight relation seen in the SPARC data does not require new
fundamental physics (such as self-interacting dark matter, modified gravity, or
MOND).  Instead, the simple combination of dissipational collapse with angular
momentum conservation produces the RAR. This means that even galaxies where SN
fail to regulate their star formation (or even where SN are omitted altogether)
fall on the same observed relation.  This was the first published demonstration
that $\Lambda CDM$ could produce the RAR, using untuned simulations made prior
to the publication of \citet{McGaugh2016}.

This thesis has shown that, with correct modelling, SN feedback can regulate the
growth of galaxies up to the mass of the Milky Way, but no further.  In galaxies
below this mass, SN efficiently drive galactic winds, primarily at high
redshift, which act to limit the availability of low angular momentum gas,
slowing the formation of stars and the stellar bulge. Without the feedback from
SMBHs, galaxies more massive than ours form far too many stars, primarily in a
concentrated central bulge.  The transition from SN to AGN feedback occurs due
to a rapid drop in the effectiveness of SN-driven outflows.  Finally, we showed
that independent of the detailed internal processes of galaxy evolution, the
acceleration relation observed by \citet{McGaugh2016} is simply a natural
consequence of galaxy formation, removing what could have been potential tension
with $\Lambda CDM$ cosmology.

\section{The Impact and Context of these Results}
\subsection{How Clustered Supernovae Heat The ISM}
Supernovae provide an enormous source of energy.  Simple accounting shows that
the energy released from a stellar population can eject more gas from a galaxy
disc than the mass of the population itself, if that energy couples to enough
mass before radiating away \citep{Larson1974}.  Radiative losses may be
significant if SN detonate in a sufficiently dense environment,
with cooling times as short as a few thousand years \citep{Chevalier1974}.
However, when massive stars form in clusters, bubbles of shocked stellar wind
and SN ejecta merge to form a superbubble.  These superbubbles have
significantly longer cooling times \citep{MacLow1988}, and can deposit an order
of magnitude more energy and momentum into the ISM before cooling
\citep{Gentry2017}.  How these superbubbles drive the 
outflows that regulate the formation of stars within the galactic disc depends
on the mass loading of these outflows.  This global outflow mass loading
depends heavily on the local mass loading: how much mass is mixed into a
superbubble before it vents from the galaxy.

The question of local mass loading has been explored before in simulation.
\citet{Thacker2000} presented a survey of methods for depositing feedback
energy.  This work showed that choosing to deposit feedback energy into single
resolution elements vs. smoothing it over multiple elements had a significant
impact on whether that feedback energy could couple to the ISM without radiating
away.  \citet{Thacker2000} advocated the use of a brief adiabatic period,
similar to the delayed cooling model used in \citet{Stinson2006} and
\citet{Rosdahl2016} to prevent overcooling when feedback energy was smoothed.  A
more recent study by \citet{DallaVecchia2012} directly examined the question of
local mass loading with a model that used the temperature of post-feedback
gas as a single tunable parameter.  This is effectively the same as choosing
the amount of mass that each SN heats (the local mass loading).
They found that lower mass loadings result in hotter post-feedback gas, with
longer cooling times, and subsequently much more effective SN feedback.
The \citet{Keller2014} superbubble model is the first model to actually
incorporate a physical mechanism for determining this mass loading.  Previous
feedback models have been designed primarily to avoid the overcooling that can
occur when this mass loading is too high: superbubble feedback instead attempts
to calculate the {\it correct} local mass loading, and allows the
feedback-heated gas to cool with realistic temperatures and densities.  The
amount of material heated by feedback is as important as the energy
available from feedback.  On a galactic scale, too-high mass loadings may mean that
more material is ejected from the disc, but at lower overall temperatures and
velocities.  This material will re-accrete on shorter timescales than hotter
ejecta, increasing the star formation rate.  Conversely, too-low mass loading
will give hot, high-entropy outflows that leave the disc efficiently and
re-accrete much later (if at all).  As these outflows contain little
mass, the ISM that remains will still be able to form stars effectively, again
increasing the star formation rate.  The results we presented here show that
thermal evaporation gives a mass-loading rate between these two extremes,
allowing effective outflows for galaxies up to $M_{vir}\sim10^{12}\Msun$.

As we showed in \citet{Keller2015}, including thermal evaporation in simulations
of a Milky Way-like galaxy drastically changes the effectiveness of SN-driven
galactic winds.  Starving the galaxy of gas (either through ejection of accreted
gas, or prevention of accretion) is the only effective mechanism for regulating
star formation on cosmic timescales \citep{Keres2009}.  To prevent the formation
of bulges larger than the ones seen in Milky Way-like galaxies, bulge-forming
(low-angular momentum) gas must be removed from the ISM.  As SN-driven winds are
most effective at high redshift, when this material is mostly accreted, the
outflows driven from the galaxy preferentially remove bulge-forming gas.  A
similar result was seen in simulations of dwarf galaxies \citep{Governato2010,
Brook2011}, but \citet{Keller2015} is the first work to show that bulgeless disc
galaxies similar to the Milky Way can be produced through the same mechanism,
and connect this with the regulation of star formation.  An important conclusion
of this study is that much of the past work
\citep{Stinson2013,Agertz2013,Hopkins2014} showing that SN feedback was unable
to regulate star formation without the addition of ``early feedback'' from
radiation pressure, stellar winds, or photoionization was likely underestimating
the effectiveness of SN.  Since many of these studies used simpler, more
numerically-sensitive feedback models that also omitted thermal evaporation,
they were unable to effectively drive sufficient mass into hot, high-entropy gas
that could leave the galaxy.  This study tells us that the disc-dominant
morphology and low stellar fraction in galaxies like our own all arise from the
same mechanism:  heavily mass loaded, high redshift winds.

\subsection{The Challenge of Modelling Feedback}
Developing a versatile, accurate, resolution-insensitive model for SN feedback
has been a challenge in galaxy simulations for decades
\citep{Katz1992,Rosdahl2016}.  While a number of recent models have been
developed that are able to produce galaxies which match some observables, all of
these models have relied on purely numerical features to enhance the
effectiveness of feedback and counter the losses due to limited resolution.  The
\citet{Keller2014} model presented in Chapter 2 is the first model of its kind
that uses a first-principles approach to prevent overcooling.  Beyond merely
providing ``sufficient'' feedback, the superbubble model also incorporates
physics that previously have been omitted.  Thermal evaporation is a key process
for mixing cold ISM with hot feedback ejecta.  Prior to \citet{Keller2014}, this
process was completely ignored, and the local mass loading of feedback was set
purely as a numerical parameter.  \citet{Rosdahl2016} illustrated this clearly,
in a recent comparison project.  This paper implemented 5 popular feedback
algorithms in the {\sc RAMSES} code: simple thermal dumps; a delayed cooling
model similar to \citet{Agertz2013}; a kinetic model similar to
\citet{DallaVecchia2008}; a stochastic thermal model based on
\citet{DallaVecchia2012}; and a hybrid model from \citet{Kimm2015}.  This
allowed for a controlled test with isolated dwarf and Milky Way analog galaxies.
Each of these models, with ``fiducial'' parameter choices, produced star
formation rates that varied by nearly an order of magnitude in the
\citet{Kennicutt1998} star formation relation.  On top of that, the galaxies
showed marked morphological differences and over 3 orders of magnitude of
variation in the average wind mass loadings.  As we showed in
\citet{Keller2015}, these winds are critical to the formation of galaxies like
the Milky Way.  In addition to the inter-model variation, \citet{Rosdahl2016}
also found that the numerical free parameters for each model had a large impact
on the resulting galaxy's star formation rate, morphology, and outflow
properties.  By building a more conservative, physically motivated model where
the local mass loading of SN feedback is set by thermal evaporation, rather than
numerically, we sidestep many of these issues.

With previous models for stellar feedback, both the qualitative behaviour as
well as the precise, quantitative effects have been heavily determined by
poorly constrained numerical parameters.  This has made interpreting the results
of simulations difficult, as it is unclear which effects are a result of
the underlying physics and which are simply a result of the choice of parameters
used for the simulation.  Physics-driven, parameter-free models remove this
ambiguity, and allow us to produce results with greater confidence.

\subsection{Feedback Beyond Supernovae}
Subgrid models are essential for including feedback from other sources as well.
Non-supernovae mechanisms for launching outflows and regulating star formation
are now frequently studied with numerical simulations.  Radiation pressure has been
examined in a number of studies \citep{Hopkins2014,Roskar2014,Agertz2015} using
fairly simple treatments, without full radiative transport.  These simulations
have included various models for feedback from SN and stellar winds, and have
not yet converged on a consensus regarding the effects of radiation.  However,
both \citet{Roskar2014} and \citet{Agertz2015} found that the optical depths
required for radiation pressure to drive effective outflows produce unrealistic
galaxy morphologies.  The effects of cosmic rays are beginning to be examined in
simulations with \citep{Girichidis2015} and without
\citep{Jubelgas2008,Booth2013} treatments for magnetohydrodynamics.  Most
simulations have found that cosmic rays are unable to drive sufficiently
mass-loaded outflows from Milky Way-like galaxies, but may be important both for
dwarf galaxies and launching the cold, dusty phases of outflows in larger
galaxies.  With a better constrained SN model, these mechanisms can now be
re-examined with greater control, allowing SN and other effects to be
disentangled.

It has been suggested by numerous studies \citep{Benson2003,Bower2006} that AGN
feedback is critical to shaping the galaxy luminosity function.  Most
observations \citep{Kauffmann2003b} show that strong AGN are hosted in galaxies
with halo masses slightly greater than $\sim10^{12}\;M_\odot$.  By simulating a
collection of galaxies that bracket this mass, but omitting the effects of AGN
feedback, we were able to show that SN alone break down as a mechanism of
regulating star formation in these massive galaxies.  Importantly, this also
results in large angular momentum transport for gas within the disc, funneling
material to the central regions where SMBHs reside.  This, combined with the
failure of SN-driven galactic winds, suggests that SN are actually the mechanism
that determines where powerful AGN occur, a mechanism that has begun to be studied in
\citet{Bower2016}.  Uncertainties in the unresolved ISM are unlikely to
significantly change the characteristic mass that ``needs'' AGN for regulation.
As we showed in Chapter 2, the strongly self-limiting nature of thermal
conduction makes the temperature of feedback-heated gas insensitive to unresolved
structure in both gas density and magnetic fields.

\subsection{Baryonic Physics \& Feedback in Tests of Cosmology}
Chapter 5 presented the first simulation result to reproduce the
\citet{McGaugh2016} RAR.  A later study \citep{Ludlow2016} used data from the
EAGLE \citep{Schaye2015} and APOSTLE \citep{Sawala2016} simulations and
confirmed both the fit to the RAR, and the insensitivity to feedback that we
found.  Earlier, a semi-empirical model used by \citet{DiCintio2016} showed
that a related observation, the mass-discrepancy acceleration relation, could be
produced simply as a result of other pre-existing scaling relations.  Taken
together, these results should seriously temper attempts to fit/explain the RAR
using new, exotic physics.  If conventional baryonic physics produces the same
relation in a $\Lambda CDM$ cosmology, there is little need to turn to modified
gravity \citep{Moffat2016,Verlinde2016,Burrage2016} to explain it.

Tests of cosmology have been moving closer to scales where the effects of
baryonic physics are becoming more pronounced.  Comparisons between the
predictions of collisionless structure formation have found that $CDM$ produces
far more dwarf galaxies than those we observe in the local group
\citep{Klypin1999,Boylan-Kolchin2011}. Observations of the dwarves we do see
shows their density profiles have flattened ``cores'', suggesting they are
depleted in dark matter within the inner few kpc \citep{Moore1994,Simon2005}.
While alternative forms of dark matter may reproduce these observations
\citep{Spergel2000,Avila-Reese2001}, it is notable that what appear to be
tension between $CDM$ simulations and the observations is actually tension
between {\it collisionless} $CDM$ simulations and the observations.  In dwarf
galaxies, the effects of stellar feedback are extremely important, with a
handful of supernovae capable of ejecting the entire ISM.  When these effects
are taken into account, much of the apparent tension disappears
\citep{Mashchenko2006,Pontzen2012,Sawala2016}.  Without well-constrained subgrid
models, like the one presented here, we will be unable to use the wealth of
information available from galaxies to test new theories of gravity or
alternative cosmologies, or constrain the effects of known physics: AGN, cosmic
rays, MHD effects, etc.  On individual galaxy scales, $\Lambda CDM$ predictions
{\it must} include the details of baryonic effects.  Chief among those effects
is feedback.

\section{Further Questions \& Future Directions}
The results presented in this thesis have greatly clarified the impact of SN
feedback, but also raised some questions.  While we now know how SN act to
regulate the growth of stars in galaxies up to the mass of the Milky Way,
details about the transition to AGN regulation at high mass are still unclear.
The outflows that shape the galaxy contain detailed, multiphase structure that
is unresolved in our simulations.  How exactly mass is entrained within these
outflows and how that material is recycled back into the disc will require
further study.  Additionally, the small sample of 18 galaxies we studied in
chapters 4 and 5 are a tiny fraction of the range of mass, morphology and
environment in real galaxies.  How does superbubble feedback work to shape the
scaling relations seen in galaxies both more and less massive than the ones we
have seen here?  The results of chapter 4 suggest that feedback from AGN is a
critical process for more massive galaxies.  How does this mechanism work in
tandem with SN to regulate the formation of massive galaxies?

\subsection{Launching Multiphase Galactic Winds}
One of the biggest questions raised in chapter 4 was the origin of the scaling
relation between outflow mass loadings and disc mass.  Observations of galactic
winds tell us that they are multiphase \citep{Stark1984,Wakker1997}, containing
ionized, neutral, and molecular gas.  This structure is unresolved in our
simulations, and may ultimately be important to the cooling rates and recycling
times for these outflows.  Complete, detailed observations of outflowing
material in the CGM are difficult. This gas tends to be too cool and diffuse to
be a strong emitter of X-rays, and absorption line studies require the lucky
coincidence of a background quasar \citep{Weiner2009}.  The presence of cold
clouds in particular is a significant problem for theories of galactic wind
launching.  Analytic calculations and high resolution simulations suggest that
these clouds should be disrupted and mixed on timescales far shorter than is
required to ballistically launch them into the CGM.  In light of this, why do we
see these clouds?  They may be formed in-situ, or may be launched by the
fragmentation of superbubble shells, already accelerated to high velocities
during the momentum-conserving snowplow phase of the bubble's growth
\citep{Lagos2013}.  These processes all take place on length scales far below
what is resolvable in the cosmological simulations we have presented here.  High
resolution simulations of superbubble breakout from the ISM will be able to tell
us what indeed is the origin of these clouds.  Understanding the details of
the multiphase CGM is key to understanding how galaxies accrete and expel gas,
as thermal instability in the CGM (which depends heavily on its density,
temperature, and metallicity) may be the mechanism by which spiral galaxies
today acquire the gas required to form new stars \citep{Marasco2012}.

A related question to the detailed, large-scale breakout of galactic winds from
the ISM is the detailed, small scale breakout of SN from their natal
environments.  Studies have already been published examining the destruction of
molecular clouds by stellar winds and ionizing radiation
\citep{Murray2010,Walch2012}.  Only a handful have focused on how SN-heated gas
can escape from molecular clouds \citep{Rogers2013}, and these have looked at
very controlled, simple initial clouds.  The transition between individual
SN in a molecular cloud and the growth of a new superbubble is 
important to determining the maximum efficiency of SN, and how much 
of the $10^{51}\;\rm{erg}$ released per SN is lost before a superbubble can begin
to form.  This, like the evolution of cold clumps in galactic winds, will
require high-resolution simulations of star formation within realistic molecular
clouds.  Studies have begun \citep{Dobbs2015} to study molecular clouds selected
from global disc simulations and re-simulated at higher resolution.  Combining
these kinds of simulations with models for small-scale star formation, feedback
from stellar winds and massive stars, as well as thermal conduction and
evaporation will allow us to determine a more accurate energy budget for stellar
feedback.

\subsection{Scaling Relations in Galaxy Populations}
While the MUGS2 sample has been simulated with a robust, physically motivated
model for SN feedback, its sample size is small (only 18 galaxies), and covers a
limited range of halo masses ($\sim 4\times10^{11}-2\times10^{12}\;M_\odot$).
Modern large volume simulations like Illustris \citep{Vogelsberger2014b} and
EAGLE \citep{Schaye2015} contain $>10^6$ individual galaxies, spanning a range
of masses that include both galaxies lighter and more massive than those we have
studied here.  These larger sample sizes are essential to understanding how
whole populations of galaxies evolved from the earliest linear structures.
While these simulations contain models for star formation, feedback from both
stars and AGN, as well as radiative cooling, these have all been fine-tuned to
fit observed scaling relations.  This limits the ability of these simulations to
explain the origins of the observations they have been tuned to fit.  The
superbubble model contains no free/tuned parameters, and accurately follows the
evolution of feedback-heated gas as it evaporates cool material that surrounds
it.  It is the only feedback model currently in existence that captures the
important physical process of thermal conduction and evaporation.  It will allow
us to build a statistically significant sample of physically realistic galaxies.
With first-principles, physically motivated models for the unresolved physics in
our simulations, we will be able to study the physics involved in shaping
both individual galaxies and the populations that we observe.

\subsection{Feedback from AGN}
While observational evidence abounds for the existence of SMBHs
\citep{Kormendy2013}, and the impact of the AGN they power
\citep{Veilleux2005}, there is still great uncertainty as to how they regulate
their own growth, and the growth of the galaxies they live in.
\citet{Keller2016a} showed evidence that a transition must occur between
galaxies regulated by SN feedback to AGN feedback as they grow heavier than the
Milky Way.  How this handoff occurs remains widely unexplored.  The superbubble
feedback model is agnostic to the source of heating: it simply captures the
effect of evaporation between unresolved hot and cold gas.  It's therefore
equally well-suited to model the effect of heating from AGN as it is for heating
from SN.  Using superbubble feedback, along with new models for SMBH growth
\citep{Hopkins2010} and migration \citep{Tremmel2015}, we will be able to
simulate the growth of SMBHs in galaxies with greater confidence, using
physically motivated feedback and accretion models.  This will allow us to study
the full gamut of galaxies that exist, and study how the tight scaling relations
between stellar bulges and SMBHs might arise.

\section{Final Thoughts}
Less than a century ago, even the most basic nature of external galaxies was
contentious.  Prominent astronomers argued ``spiral nebulae'' were members of
our own galaxy, the Milky Way.  Today, we stand on the shoulders of legion
observers, theorists, and simulators in constructing a self-consistent theory of
galaxy formation.  While these past generations have painted a picture of galaxy
formation as gravitational collapse through hierarchical, bottom-up merging,
details remain unclear.  Perhaps the most puzzling \& important question is the
inefficiency of star formation at all scales.  At best, a galaxy may convert a
third of its baryons into stars, with most galaxies forming stars far more
ineffectively.  This thesis has explored one mechanism for this inefficiency,
feedback by SN .  We have shown that thermal conduction is critical to
evaporating the star-forming ISM into hot outflows that starve young,
high-redshift galaxies of star formation fuel.  This mechanism is effective in
galaxies up to the mass of our own; it breaks down in more massive galaxies.
This tells us two important facts about how galaxies evolve.  First, it tells us
that SN alone may regulate star formation in galaxies with $M_{vir} <
10^{12}\Msun$.  Second, it tells us that SN {\it cannot} be the only source of
energy that regulates the formation of massive galaxies.  While many questions
remain about the processes that govern the life cycle of galaxies, the results
shown here provide a solid foundation for understanding one of the most
important processes: how SN regulate galaxy formation.

\bibliographystyle{mnras}
\bibliography{library}